---
title: "Data wrangling and reverse engineering"
output: pdf_document
---

This is a hopefully straightforward set of tasks. As part of the homework you will find the file RADxUP_CDE_Tracking_full.csv

This is a file saved from a fairly complicated set of questionnaires from a real set of projects. Each row is a project, and each column is a item from the questionnaire. In order to evaluate the responses, I created a simple heatmap of some of the answers. The data is from the first 25 rows and columns 12-42. The heatmap I made is also attached in the homework, called newplot.png. 

Task 1. Create the code necessary to load the file and prepare it for heatmaply. Insert the code below. 

```{r}
rad_data <- read.csv("../Visualization_HW4/RADxUP_CDE_Tracking_full.csv")
dim(rad_data)
```
 
How many rows and columns are in the file?  
Number of rows: `r dim(rad_data)[1]`

Number of columns: `r dim(rad_data)[2]`

Task 2. Create a simple heatmap using heatmaply that replicates/approximates the heatmap I put in the homework. Insert your code and heatmap below. Note that you will need to adjust the fontsize or else the plotsize to make the project names legible. 

```{r}
library(arsenal)
TNE <- c("Johns Hopkins University", "Washington State University", " Montana State University - Bozeman")

# rad_data %>% filter(Tribal.Nation.exception == TRUE)
map_data_NA <- rad_data %>%  filter(Awardee.Name %nin% TNE)

heatmaply(
  map_data_NA,
  xlab = "Features",
  ylab = "Cars",
  main = "Data"
)

# heatmaply_na(
#   rad_data[2:ncol(rad_data)],
#   showticklabels = c(TRUE, FALSE),
#   k_col = 3,
#   k_row = 3,
#   # fontsize_row = 6,
#   # fontsize_col = 6
# )

```

Task 3. Create a heatmap and adjust the options to try and show all the projects in a simple map. Insert your code below:


```{r}

```

Task 4. Now start mapping additional columns. Create a series of plots (or a single legible plot if you can!) for all the columns. You will need to shorten/edit the column names at some point in the file. Try and keep the column names informative/meaningful. Insert your code below


```{r}

```

Task 5. Come up with another way to plot these data in a heatmap. Feel free to explore other heatmap plotting tools in R Studio. Insert your code below:

```{r}

```

Task 6. While heatmaps are very compact and an efficient way to show this kind of data, experiment with some other methods. Can you find another way to plot/visualize these data? Include your code below:

```{r}

```

